---
title: "biocUniTools"
format:
  html:
    toc: false
sidebar: false
editor: visual
engine: knitr
vignette: >
  %\VignetteIndexEntry{biocUniTools}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

biocUniTools aims to make it easier to work with Bioconductor package data in
R Universe. It can retrieve raw data for a universe or attempt to match jobs
to binaries by OS and arch. The package also contains small helper functions to
manage repository clean up when maintaining a package repository that's
replenished with new binaries from R Universe.

```{r install_universe, echo = FALSE, message = FALSE}
library(dplyr)
library(biocUniTools)
library(reactable)
```

## R Universe API

Check a package like [lefser](https://bioc.r-universe.dev/lefser) in R Universe. The R Universe API has the same data available programmatically at
https://bioc.r-universe.dev/api/packages/lefser.

Data comes from the `DESCRIPTION` file and GitHub. In particular, we're 
interested in the `_jobs`, `_binaries`, and `_buildurl` when considering whether
to propagate a binary created in R Universe. Notice it has jobs for 
`bioc-checks`, builds on multiple versions of R, and the `_jobs_config` doesn't
exactly match a `_binaries`'s `os`.

## lefser in Bioconductor 3.23 via the R Universe API

This is a truncated version of parts of the API.

```{json lefser-in-3.23-on-2026047}
{
  "_id": "69cf91f20c54e422bf1d50de",
  "Package": "lefser",
  "Type": "Package",
  "Title": "R implementation of the LEfSE method for microbiome biomarker\ndiscovery",
  "_jobs": [
    {
      "job": 69829927166,
      "time": 244,
      "config": "bioc-checks",
      "r": "4.5.3",
      "check": "NOTE",
      "artifact": "6257774022"
    },
    {
      "job": 69829927169,
      "time": 379,
      "config": "linux-devel-x86_64",
      "r": "4.6.0",
      "check": "OK",
      "artifact": "6257796829"
    },
    {
      "job": 69829927167,
      "time": 449,
      "config": "linux-release-x86_64",
      "r": "4.5.3",
      "check": "OK",
      "artifact": "6257808198"
    },
    {
      "job": 69829927184,
      "time": 252,
      "config": "macos-oldrel-arm64",
      "r": "4.5.3",
      "check": "ERROR",
      "artifact": "6258002602"
    },
    {
      "job": 69829927171,
      "time": 282,
      "config": "macos-release-arm64",
      "r": "4.6.0",
      "check": "ERROR",
      "artifact": "6258007467"
    },
    {
      "job": 69829465206,
      "time": 327,
      "config": "source",
      "r": "4.5.3",
      "check": "OK",
      "artifact": "6257731498"
    },
    {
      "job": 69829927173,
      "time": 230,
      "config": "wasm-release",
      "r": "4.5.1",
      "check": "OK",
      "artifact": "6257771337"
    },
    {
      "job": 69829927228,
      "time": 615,
      "config": "windows-devel",
      "r": "4.7.0",
      "check": "ERROR",
      "artifact": "6257837183"
    },
    {
      "job": 69829927188,
      "time": 517,
      "config": "windows-oldrel",
      "r": "4.5.3",
      "check": "ERROR",
      "artifact": "6257820249"
    },
    {
      "job": 69829927225,
      "time": 643,
      "config": "windows-release",
      "r": "4.6.0",
      "check": "ERROR",
      "artifact": "6257841770"
    }
  ],
  "_bioccheck": {
    "error": 0,
    "warning": 0,
    "note": 8
  },
  "_buildurl": "https://github.com/r-universe/bioc/actions/runs/23941765469",
  "_binaries": [
    {
      "r": "4.6.0",
      "os": "linux",
      "version": "1.21.7",
      "date": "2026-04-03T09:45:48.000Z",
      "distro": "noble",
      "commit": "333fd2411085100ddc70fdbffbf1cbebb735df8d",
      "fileid": "0523834cf0d6f2942ec27c4cebd0ce8fd20f317967eb7c69672abf0642a95426",
      "status": "success",
      "check": "OK",
      "buildurl": "https://github.com/r-universe/bioc/actions/runs/23941765469"
    },
    {
      "r": "4.5.3",
      "os": "linux",
      "version": "1.21.7",
      "date": "2026-04-03T09:46:50.000Z",
      "distro": "noble",
      "commit": "333fd2411085100ddc70fdbffbf1cbebb735df8d",
      "fileid": "2906aa666beaae1e887c356eb68fcf5cca2eabff42aefea05399215fade84680",
      "status": "success",
      "check": "OK",
      "buildurl": "https://github.com/r-universe/bioc/actions/runs/23941765469"
    },
    {
      "r": "4.5.3",
      "os": "mac",
      "version": "1.21.7",
      "date": "2026-04-03T10:06:49.000Z",
      "commit": "333fd2411085100ddc70fdbffbf1cbebb735df8d",
      "fileid": "d9e157f9bb0160ad67bbb54a911a6c921e7c4e1f25c856afd6554571603d4991",
      "status": "failure",
      "check": "ERROR",
      "buildurl": "https://github.com/r-universe/bioc/actions/runs/23941765469"
    },
    {
      "r": "4.6.0",
      "os": "mac",
      "version": "1.21.7",
      "date": "2026-04-03T10:07:02.000Z",
      "commit": "333fd2411085100ddc70fdbffbf1cbebb735df8d",
      "fileid": "e4083909ae44d1d8317c445f939e547b5f1d314bb0d0e68982ae31eaee9e4fbe",
      "status": "failure",
      "check": "ERROR",
      "buildurl": "https://github.com/r-universe/bioc/actions/runs/23941765469"
    }
  ]
}
```

Binaries that work on any arch of an OS are universal binaries. Universal
binaries are built on the x86_64 linux whereas for macosx, it's built on
arm64. However, if your package must compile C code, it needs a binary that's
arch-specific so the R Universe API will include additional fields for these
binaries. This information is important to construct the repository path. Note:
the `arch` field in `rtracklayer`'s `_binaries` for macosx.

```{json rtracklayer-binaries-for-mac-aarch}
}
  "_binaries": [
    {
      "r": "4.6.0",
      "os": "mac",
      "version": "1.71.3",
      "date": "2026-03-16T06:33:54.000Z",
      "arch": "aarch64",
      "commit": "d9b23f963c621e4aa8e5e52c6e4da50d5d213e1a",
      "fileid": "ac6b65936a87f30994737a041f03c22f0d85c82b0a22b3a31b3e01c98ab3d5e1",
      "status": "success",
      "check": "WARNING",
      "buildurl": "https://github.com/r-universe/bioc/actions/runs/23130760393"
    }
}
```

## Using biocUniTools

Let's imagine we want to create a repository of R Universe binaries for macosx
arm64. We'll use `biocUniTools` to align `_jobs` and `_binaries` and filter
for candidate packages that match our criteria and show to clean up our
repository for future maintenance.

Get corresponding R Universe information for a Bioconductor branch.

```{r get-uni-for-bioc-branch}
bioc_info <- get_uni_for_bioc_version("3.23")
bioc_info
```

If we wanted to create a report of each package's status, we could use `get_jobs`,
which attempts to align jobs and binaries where possible. This function
creates a few extra columns to help sort differences in these data points, a
`JobUrl`to the GitHub Action job, and `Artifact`, the url for the binary in
R Universe.

::: column-screen-inset

```{r get_jobs}
jobs <- get_jobs(bioc_info$ru_uni, bioc_info$bioc_branch, bioc_info$r_version,
                 bioc_info$bioc_version)

jobs |>
    dplyr::filter(Package %in% c("lefser", "rtracklayer")) |>
    dplyr::select(Package, `_jobs_r_xy`, `_jobs_type`, `_jobs_arch`,
                  `_binaries_arch`, `_jobs_os_`, `_binaries_os_`,
                  `_jobs_check`, `_binaries_check`)
```
:::

Use `get_candidates` to find candidates binaries if their vignettes have passed
check and the commit is the same in the BBS. We could use `curl::multi_download`
to download all binaries into our repository.
::: column-screen-inset
```{r get_candidates}
candidates <- get_candidates(bioc_info$ru_uni, bioc_info$bioc_branch,
                             bioc_info$r_version, bioc_info$bioc_version,
                             "macosx", "arm64", vignettes = TRUE,
                             commit = TRUE)
candidates |>
    dplyr::filter(Package %in% c("lefser", "rtracklayer")) |>
    dplyr::select(Package, `_jobs_r_xy`, `_jobs_type`, `_jobs_arch`,
                  `_jobs_os_`, `_jobs_check`, `_binaries_check`, JobUrl,
                  Artifact)
```
:::

`biocUniTools` also has `remove_old_binaries` to clean up older versions of
binaries after updating the repository with R Universe binaries. See
`inst/scripts/harvest.R` for a more complete example.

```{r remove-old-binaries, eval = FALSE}
os <- "macosx"
arch <- "arm64"
repo_root <- "~/tmp/PACKAGES/3.23/bioc"
# The R 4.6 macosx path uses sonoma for arm64
repo_path <- get_repository_path(repo_root, bioc_info$r_version, os, "sonoma",
                                 arch)
downloaded <- curl::multi_download(candidates$Artifact, destfiles = repo_path)
remove_old_binaries(repo_root, bioc_info$r_version, os, arch)
```

```{r sessionInfo}
sessionInfo()
```
