---
title: "biocUniTools"
format: revealjs
editor: visual
execute:
  echo: true
---

## Motivation

-   [Bioconductor moving to R Universe](https://blog.bioconductor.org/posts/2026-04-08-r-universe-collaboration/)
-   Support for working with Bioconductor data in the R Universe API
    -   Understanding overall health
    -   Binary collection
    -   Propagation criteria

## binaries and propagation +

-   binary: prebuilt packages for macosx and windows users that make it easy to
    install Bioconductor packages because compilation is sometimes hard.
-   propagation: based a set of criteria, make a new version of the package available
    -   Criteria on BBS: passes `R CMD INSTALL`, `R CMD build`,
        `R CMD check`, `R CMD buildbin` (this makes the actual binary)
    -   Criteria in RU: `R CMD check`, vignettes passing, binaries check pass,
        binaries status pass, job check pass, and for now matching commit in
        BBS and RU

## RU API

Combines data from package `DESCRIPTION` files and Github Action workflow runs.

Different

-   terminologies: BiocCheck vs `bioc-checks`, WARNINGS vs WARNING
-   abbreviations: macosx vs macos vs mac or aarch64 vs arm64
-   data: no `R CMD INSTALL`, `R CMD build` (ish), etc.

Also

-   extra fields only if needed: universal vs arch-specific builds
-   jobs and binaries separated

## Universal and arch-specific binaries + 

Needed to compile, C for example, in an R package.

Universal binary building

os | arch | universal builder?
---+------+-------------------
linux | x86_64 | Yes
linux | aarch64 | No
macosx | x86_64 | No
macosx | arm64 | Yes

## Comparison: lefser

```{json lefser, echo = TRUE}
{
  "_binaries": [
    {
      "r": "4.6.0",
      "os": "mac",
      "version": "1.21.7",
      "date": "2026-04-03T10:07:02.000Z",
      "commit": "333fd2411085100ddc70fdbffbf1cbebb735df8d",
      "fileid": "e4083909ae44d1d8317c445f939e547b5f1d314bb0d0e68982ae31eaee9e4fbe",
      "status": "failure",
      "check": "ERROR",
      "buildurl": "https://github.com/r-universe/bioc/actions/runs/23941765469"
    }
  ]
}
```
[lefser](https://bioc.r-universe.dev/lefser)

## Comparison: rtracklayer

```{json rtracklayer-binaries-for-mac-aarch, echo = TRUE}
}
  "_binaries": [
    {
      "r": "4.6.0",
      "os": "mac",
      "version": "1.71.3",
      "date": "2026-03-16T06:33:54.000Z",
      "arch": "aarch64",
      "commit": "d9b23f963c621e4aa8e5e52c6e4da50d5d213e1a",
      "fileid": "ac6b65936a87f30994737a041f03c22f0d85c82b0a22b3a31b3e01c98ab3d5e1",
      "status": "success",
      "check": "WARNING",
      "buildurl": "https://github.com/r-universe/bioc/actions/runs/23130760393"
    }
  ]
}
```
[rtracklayer](https://bioc.r-universe.dev/rtracklayer)

## biocUniTools

Makes it easier to work with Bioconductor data in RU by

-   dealing with differences
-   match `_binaries` and `_jobs` where possible
-   get candidate packages based on criteria
-   helpers to generate urls to binary artifacts and job build urls on Github

## Examples

Get corresponding R Universe information for a Bioconductor branch.

```{r get-uni-for-bioc-branch}
library(biocUniTools)

bioc_info <- get_uni_for_bioc_version("devel")
bioc_info
```

## `get_jobs` to create a report

::: column-screen-inset

```{r get_jobs}
jobs <- get_jobs(bioc_info$ru_uni, bioc_info$bioc_branch, bioc_info$r_version,
                 bioc_info$bioc_version)

jobs |>
    dplyr::filter(Package %in% c("lefser", "rtracklayer")) |>
    dplyr::select(Package, `_jobs_r_xy`, `_jobs_type`, `_jobs_arch`,
                  `_binaries_arch`, `_jobs_os_`, `_binaries_os_`,
                  `_jobs_check`, `_binaries_check`)
```
:::

## Use case: create a macosx arm64 binary repository

Use `get_candidates` to find candidates binaries if their vignettes have passed
check and the commit is the same in the BBS. We could use `curl::multi_download`
to download all binaries into our repository.

::: column-screen-inset
```{r get_candidates}
candidates <- get_candidates(bioc_info$ru_uni, bioc_info$bioc_branch,
                             bioc_info$r_version, bioc_info$bioc_version,
                             "macosx", "arm64", vignettes = TRUE,
                             commit = TRUE)
candidates |>
    dplyr::filter(Package %in% c("lefser", "rtracklayer")) |>
    dplyr::select(Package, `_jobs_r_xy`, `_jobs_type`, `_jobs_arch`,
                  `_jobs_os_`, `_jobs_check`, `_binaries_check`, JobUrl,
                  Artifact)
```
:::

## Clean up with `remove_old_binaries`

Clean up older versions of binaries after updating the repository with RU
binaries. See `inst/scripts/harvest.R` for a more complete example.

```{r remove-old-binaries, eval = FALSE}
os <- "macosx"
arch <- "arm64"
repo_root <- paste0("~/tmp/PACKAGES/", bioc_info$bioc_version, "/bioc")
# The R 4.6 macosx path uses sonoma for arm64
repo_path <- get_repository_path(repo_root, bioc_info$r_version, os, "sonoma",
                                 arch)
downloaded <- curl::multi_download(candidates$Artifact, destfiles = repo_path)
remove_old_binaries(repo_root, bioc_info$r_version, os, arch)
```

## `sessionInfo`

```{r sessionInfo}
sessionInfo()
```
