---
title: "Bioconductor in R Universe: Linux and Vignettes"
format:
  html:
    code-fold: true
editor: visual
engine: knitr
---

```{r install_universe, echo = FALSE, message = FALSE}
library(BiocPkgTools)
library(dplyr)
library(biocUniTools)
library(httr2)
library(reactable)
library(purrr)
```

```{r get_buildurl, echo = FALSE}
# When a package fails early in the build, we don't get a _buildurl but we can
# construct it
get_buildurl <- function(pkg, uni) {
    buildurl <- tryCatch({
        u <-paste0("https://", uni, ".r-universe.dev/api/packages/", pkg)
        pkglist <- httr2::request(u) |>
            httr2::req_timeout(10) |>  # timeout after 10 seconds
            httr2::req_perform() |>
            httr2::resp_body_json()
        pkglist$`_buildurl`
        
    }, error = function(e) {
            if (is.null(e$resp)) return(NA_character_)
            body <- httr2::resp_body_string(e$resp)
            buildurl <- stringr::str_extract(body, 
                                             paste0("https://github.com/r-universe/",
                                                    uni, "/actions/runs/[0-9]+"))
            buildurl
    })
    buildurl
}
```

```{r propagation-criteria, echo = FALSE}
propagate <- function(commit_match, deprecated, job_check, binary_check,
                      binary_status, vignettes_check) {
    !is.na(commit_match) & commit_match & 
        !is.na(deprecated) & !deprecated &
        !is.na(job_check) & (!job_check %in% c("CANCELLED", "ERROR", "FAIL")) &
        !is.na(binary_check) & (!binary_check %in% c("CANCELLED", "ERROR", "FAIL")) &
        !is.na(binary_status) & (binary_status == "success") &
        !is.na(vignettes_check) & (!vignettes_check %in% c("CANCELLED", "ERROR", "FAIL"))
}
```

The following tables attempt to describe the potential propagation status of a
package in R universe. Propagation is based on

- the same commit in both build systems
- the job building in R Universe
- the binary building in R Universe
- the vignettes building in R Universe
- package is not deprecated

Possible failure states are `CANCELLED`, `ERROR`, or `FAIL`.

## R Universe and Bioconductor 3.23

https://bioc.r-universe.dev

```{r devel-settings, echo = FALSE, message = FALSE}
r_version <- "4.6.0"
bioc_version <- "3.23"
uni <- "bioc"
uni_os_branch <- "devel"
os <- "linux"
nde <- "nebbiolo1"
ru_builds <- biocUniTools::get_jobs(uni, uni_os_branch, r_version, bioc_version)
bbs_pkgs <- BiocPkgTools::biocBuildReport(version = bioc_version,
                                            pkgType = "software")
```

::: column-screen-inset
```{r devel-data, echo = FALSE}
bbs_builds <- bbs_pkgs |>
    dplyr::filter(node == nde, stage == "checksrc") |>
    dplyr::select(pkg, version, Deprecated, git_last_commit, result) |>
    dplyr::rename(Package = pkg, Version = version,
                  `BBS Commit` = git_last_commit, Check = result) |>
    dplyr::mutate(`BBS Commit` = substr(`BBS Commit`, 1, 7)) |>
    dplyr::distinct()

filtered_ru_builds <- ru_builds |>
    dplyr::filter(`_jobs_os_` == os) |>
    biocUniTools::filter_by_arch(os)
vignettes_ru_builds <- ru_builds |>
    dplyr::filter(`_jobs_type` == "source") |>
    dplyr::select(Package, Version, `_jobs_check`, `_buildurl`) |>
    dplyr::rename(`Vignettes Check` = `_jobs_check`,
                  vignettes_buildurl = `_buildurl`)

builds <- merge(bbs_builds, filtered_ru_builds, by = c("Package", "Version"),
                all.x = TRUE)
builds <- merge(builds, vignettes_ru_builds, by = c("Package", "Version"),
                all.x = TRUE)

# Get _buildurl if not available
needs_url <- builds |>
    dplyr::filter(!Deprecated & is.na(`_buildurl`)) |>
    dplyr::distinct(Package, .keep_all = TRUE)

needs_url$`_buildurl` <- purrr::map_chr(needs_url$Package, get_buildurl, uni)

builds <- builds |>
    dplyr::rows_update(needs_url |> dplyr::select(Package, `_buildurl`),
                       by = "Package",
                       unmatched = "ignore")

condensed_builds <- builds |>
    dplyr::rename(`Job Check` = `_jobs_check`, `Binary Check` = `_binaries_check`,
                  `Build Url` = `_buildurl`, `RU Commit` = RemoteSha,
                  `Job Url` = JobUrl, `Binary Status` = `_binaries_status`) |>
    dplyr::mutate(`RU Commit` = substr(`RU Commit`, 1, 7)) |>
    dplyr::mutate(`Commit Match` = `BBS Commit` == `RU Commit`,
                  Propagate = propagate(`Commit Match`, Deprecated,
                                        `Job Check`, `Binary Check`,
                                        `Binary Status`, `Vignettes Check`)) |>
    dplyr::select(Package, Version, `BBS Commit`, Deprecated, `Commit Match`,
                  Check, `Job Check`, `Binary Check`, `Binary Status`,
                  `Job Url`, `Build Url`, `Vignettes Check`, vignettes_buildurl,
                  Propagate)

condensed_builds_display <- condensed_builds |>
    dplyr::mutate(
        Check = ifelse(!is.na(Check),
            paste0('<a href="https://bioconductor.org/checkResults/', bioc_version,
                   '/bioc-LATEST/', Package, '" target="_blank">', Check, '</a>'),
            as.character(Check)),
        `Job Check` = ifelse(!is.na(`Job Url`) & !is.na(`Job Check`),
            paste0('<a href="', `Job Url`, '" target="_blank">', `Job Check`, '</a>'),
            as.character(`Job Check`)),
        `Binary Check` = ifelse(!is.na(`Build Url`) & !is.na(`Binary Check`),
            paste0('<a href="', `Build Url`, '" target="_blank">', `Binary Check`,
                   '</a>'),
            as.character(`Binary Check`)),
        `Vignettes Check` = ifelse(!is.na(vignettes_buildurl) & !is.na(`Vignettes Check`),
            paste0('<a href="', vignettes_buildurl, '" target="_blank">',
                   `Vignettes Check`, '</a>'),
            as.character(`Vignettes Check`)),
        Propagate = ifelse(Propagate,
            '<span>&#x2705;</span>',
            '<span>\U0001F6AB</span>')
    ) |>
    dplyr::select(-`Job Url`, -`Build Url`, -vignettes_buildurl)

reactable::reactable(condensed_builds_display, searchable = TRUE, #wrap = FALSE,
                     showPageSizeOptions = TRUE, compact = TRUE,
                     columns = list(
                       Check = reactable::colDef(html = TRUE),
                       `Job Check` = reactable::colDef(html = TRUE),
                       `Binary Check` = reactable::colDef(html = TRUE),
                       `Vignettes Check` = reactable::colDef(html = TRUE),
                       Propagate = reactable::colDef(html = TRUE)

                     ),
                     defaultColDef = reactable::colDef(
                       minWidth = 150,
                       style = list(whiteSpace = "nowrap"),
                       headerStyle = list(whiteSpace = "nowrap")
                     )
)
```
:::

## R Universe and Bioconductor 3.22

https://bioc-release.r-universe.dev

```{r release-settings, echo = FALSE, message = FALSE}
r_version <- "4.5.0"
bioc_version <- "3.22"
uni_os_branch <- "release"
uni <- "bioc-release"
nde <- "nebbiolo2"
ru_builds <- biocUniTools::get_jobs(uni, uni_os_branch, r_version, bioc_version)
bbs_pkgs <- BiocPkgTools::biocBuildReport(version = bioc_version,
                                          pkgType = "software")
```

::: column-screen-inset
```{r release-data, echo = FALSE}
bbs_builds <- bbs_pkgs |>
    dplyr::filter(node == nde, stage == "checksrc") |>
    dplyr::select(pkg, version, Deprecated, git_last_commit, result) |>
    dplyr::rename(Package = pkg, Version = version,
                  `BBS Commit` = git_last_commit, Check = result) |>
    dplyr::mutate(`BBS Commit` = substr(`BBS Commit`, 1, 7)) |>
    dplyr::distinct()

filtered_ru_builds <- ru_builds |>
    dplyr::filter(`_jobs_os_` == os) |>
    biocUniTools::filter_by_arch(os)
vignettes_ru_builds <- ru_builds |>
    dplyr::filter(`_jobs_type` == "source") |>
    dplyr::select(Package, Version, `_jobs_check`, `_buildurl`) |>
    dplyr::rename(`Vignettes Check` = `_jobs_check`,
                  vignettes_buildurl = `_buildurl`)

builds <- merge(bbs_builds, filtered_ru_builds, by = c("Package", "Version"),
                all.x = TRUE)
builds <- merge(builds, vignettes_ru_builds, by = c("Package", "Version"),
                all.x = TRUE)

# Get _buildurl if not available
needs_url <- builds |>
    dplyr::filter(!Deprecated & is.na(`_buildurl`)) |>
    dplyr::distinct(Package, .keep_all = TRUE)

needs_url$`_buildurl` <- purrr::map_chr(needs_url$Package, get_buildurl, uni)

builds <- builds |>
    dplyr::rows_update(needs_url |> dplyr::select(Package, `_buildurl`),
                       by = "Package",
                       unmatched = "ignore")

condensed_builds <- builds |>
    dplyr::rename(`Job Check` = `_jobs_check`, `Binary Check` = `_binaries_check`,
                  `Build Url` = `_buildurl`, `RU Commit` = RemoteSha,
                  `Job Url` = JobUrl, `Binary Status` = `_binaries_status`) |>
    dplyr::mutate(`RU Commit` = substr(`RU Commit`, 1, 7)) |>
    dplyr::mutate(`Commit Match` = `BBS Commit` == `RU Commit`,
                  Propagate = propagate(`Commit Match`, Deprecated,
                                        `Job Check`, `Binary Check`,
                                        `Binary Status`, `Vignettes Check`)) |>
    dplyr::select(Package, Version, `BBS Commit`, Deprecated, `Commit Match`,
                  Check, `Job Check`, `Binary Check`, `Binary Status`,
                  `Job Url`, `Build Url`, `Vignettes Check`, vignettes_buildurl,
                  Propagate)

condensed_builds_display <- condensed_builds |>
    dplyr::mutate(
        Check = ifelse(!is.na(Check),
            paste0('<a href="https://bioconductor.org/checkResults/', bioc_version,
                   '/bioc-LATEST/', Package, '" target="_blank">', Check, '</a>'),
            as.character(Check)),
        `Job Check` = ifelse(!is.na(`Job Url`) & !is.na(`Job Check`),
            paste0('<a href="', `Job Url`, '" target="_blank">', `Job Check`, '</a>'),
            as.character(`Job Check`)),
        `Binary Check` = ifelse(!is.na(`Build Url`) & !is.na(`Binary Check`),
            paste0('<a href="', `Build Url`, '" target="_blank">', `Binary Check`,
                   '</a>'),
            as.character(`Binary Check`)),
        `Vignettes Check` = ifelse(!is.na(vignettes_buildurl) & !is.na(`Vignettes Check`),
            paste0('<a href="', vignettes_buildurl, '" target="_blank">',
                   `Vignettes Check`, '</a>'),
            as.character(`Vignettes Check`)),
        Propagate = ifelse(Propagate,
            '<span>&#x2705;</span>',
            '<span>\U0001F6AB</span>')
    ) |>
    dplyr::select(-`Job Url`, -`Build Url`, -vignettes_buildurl)

reactable::reactable(condensed_builds_display, searchable = TRUE, #wrap = FALSE,
                     showPageSizeOptions = TRUE, compact = TRUE,
                     columns = list(
                       Check = reactable::colDef(html = TRUE),
                       `Job Check` = reactable::colDef(html = TRUE),
                       `Binary Check` = reactable::colDef(html = TRUE),
                       `Vignettes Check` = reactable::colDef(html = TRUE),
                       Propagate = reactable::colDef(html = TRUE)

                     ),
                     defaultColDef = reactable::colDef(
                       minWidth = 150,
                       style = list(whiteSpace = "nowrap"),
                       headerStyle = list(whiteSpace = "nowrap")
                     )
)
```
:::

```{r sessionInfo}
sessionInfo()
```
